On the Synchronization Power of the Openflow Data Plane

### 2015-03-22 inband-STN.tex
updating section 5

\usepackage{amssymb}
\usepackage{comment}
\usepackage{amsmath}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsRule{*}{mps}{*}{<++>}

@@

\newcommand{\hide}[1]{}
\newcommand{\concat}[0]{\oplus}
\newcommand{\append}[0]{\frown}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\cas}{CAS}
\newcommand{\claimcheck}{check}
\newcommand{\compare}{compare}

@@

\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\subsection{The flow entry modification command (FlowMod)}

\subsubsection{The CHECK\_OVERLAP flag}

\subsection{Bundling}

@@

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which can be used for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{\claimcheck} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that explained in Section \ref{sec:background}. We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.

@@

 As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer, 
  the match part of the added flow entry represents the claimed id (id concatenated twice).
   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.

In order to check if a number is claimed we try to create the corresponding entry while setting the
flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists. If the check succeeds we delete the entry. Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$. Therefore if one is used in check after the other used in claims, a conflict is detected iff $x_1=x_2$, as expected.
Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail. Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.

@@

We use similar techniques for implementing memory write and abortIfUnequal operations.
 As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
  the match part of the added flow entry is used as the memory address and the action as the written value. Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).

@@

    		\State $cmd1\gets \textsc{abortIfUnequal}((addr, old) $
    		\State $cmd2\gets \textsc{Write}((addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}

@@

%\section{The Container Object}\label{sec:vision}
\section{Transactional Synchronization}\label{sec:main}

@@

@@

\begin{algorithm}[t]
    \caption{Policy update with only CAS}
    \label{alg:simple-update}
    \begin{algorithmic}[1]
    \Require update requirements $U$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $pid\gets$ current policy id 
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
	 			\State $CAS(pid,pid+1)$
	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\begin{algorithm}[t]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ an unused \emph{id} 
 			\State $pid\gets$ current policy id 
 			\State $claim(pid)$
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
	 			\State $unclaim(pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
 				\State $abortIfClaimed(my\_id)$
	 			\State $CAS(pid,my\_id)$
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

>>>>

\usepackage{amssymb}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{xspace}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsRule{*}{mps}{*}{<++>}

@@

\newcommand{\hide}[1]{}
\newcommand{\concat}[0]{\oplus}
\newcommand{\append}[0]{\frown}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\cas}{CAS\xspace}
\newcommand{\claimcheck}{check\xspace}
\newcommand{\compare}{compare\xspace}
\newcommand{\memwrite}{write\xspace}

@@

\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\subsection{The flow entry modification command (FlowMod)}
- Creating an entry with the same match part replaces the existing one.

\subsubsection{The CHECK\_OVERLAP flag}
- creating and entry with overlapping match (same priority) and different instructions fails.
\subsection{Bundling}
- failed command aborts the bundle.

@@

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which we later use for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{\claimcheck} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that we explain in Section \ref{sec:background}. We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.

@@

 As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, for the resource-claimer, 
  the match part of the added flow entry represents the claimed id (id concatenated twice).
   In addition the mask part of the match (controller id concatenated with all-ones) varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
    We can see that for any two resource ids $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_2=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both be used to claim and unclaim without affecting each other.

In order to check if a number is claimed we try to create the corresponding entry while setting the
flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists. If the check succeeds we delete the entry. Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$. Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.

Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail. Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.

@@

We use similar techniques for implementing memory write and \compare 
 operations.
 As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
  the match part of the added flow entry is used as the memory address and the action as the written value. Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).

@@

    		\State $cmd1\gets \textsc{\compare}((addr, old) $
    		\State $cmd2\gets \textsc{\memwrite}((addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}

@@

\hide{
%\section{The Container Object}\label{sec:vision}
\section{Transactional Synchronization}\label{sec:main}

@@

\begin{algorithm}[t]
    \caption{$\textit{try-commit}()$}
    \label{alg:commit}
    \begin{algorithmic}[1]
    \Require  current bundle id $bid$.
		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
    		\State  $cmd.bundle\_id \gets bid$
    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
    		\State send $cmd$
		\Return $cmd$
    \end{algorithmic}
\end{algorithm}
}

@@

\begin{algorithm}[t]
    \caption{Policy update with only CAS}
    \label{alg:simple-update}
    \begin{algorithmic}[1]
    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $pid\gets$ current policy id 
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
	 			\State $CAS(pid\_addr,pid,pid+1)$
	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\begin{algorithm}[t]
    \caption{Advanced policy update}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$, a predefined (memory) address for  policy id: $pid\_addr$
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ an unused \emph{id} 
 			\State $pid\gets$ current policy id 
 			\State $claim(pid)$
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
	 			\State $unclaim(pid)$
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTxn
 				\State $\claimcheck(my\_id)$
	 			\State $CAS(pid\_addr, pid,my\_id)$
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTxn
	 		\State $unclaim(pid)$
     		%\State $res \gets $ commit transaction
     	\Until{$res\neq\texttt{abort}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return
